Baseline centering problem

### 2018-12-19 paper/tangotd.tex

\paragraph{Better variant?}
I think
everything works equally well if we just learn a kernel $Z(s_1,s_2)$ such
that
\begin{equation}
\transp{\1_{s_1}}\Delta^{-1}\psi\approx \E_{s_2\sim \rho} \,Z(s_1,s_2)
\psi(s_2)
\end{equation}
and this is still linear in $\psi$. This represents the inverse Laplace
operator by
\begin{equation}
Z=\sum_{s_1,s_2}\rho(s_2)  Z(s_1,s_2)\1_{s_1}\transp{\1_{s_2}}
\end{equation}
In particular the value function is
\begin{equation}
V(s_1)=\E_{s_2\sim \rho} \, Z(s_1,s_2) R(s_2)
\end{equation}
Downside: this representation of the value function does not allow for first
computing a representation of $R$ as $\E_{s_2} R(s_2)f(s_2)$ online, then taking a dot product with
$\green$. Here we would have to learn an auxiliary network to imitate
$\E_{s_2\sim \rho} Z(s_1,s_2) R(s_2)$ (but this is a supervised learning
problem). Alternatively, policy gradient can just directly be obtained
via
\begin{equation}
\partial \pi(a|s)\E_{s'|s,a} \E_{s_2\sim \rho} \, Z(s',s_2)R(s_2)
\end{equation}
(this works even if $\rho$ is \emph{not} the invariant distribution). We
can apply this each time a new reward is observed at some point $s_2$. We
can even set another arbitrary distribution on $s_1$: this policy
gradient will maximize
the reward obtained from a starting point from that distribution.

@@

On-policy, sampling from $Z(\cdot,s_3)$ is easy: just sample one of the
recent states with geometric rates $\gamma$. This should divide the
right part of the equation by $Z(s_1,s_3)$ (up to some $\gamma$-dependent
factor). Thus we would sample two states $s_1,s_2$ along the trajectory,
sample a state $s_3$ in between on the trajectory from $s_1$, and (with
the right $\gamma$ factors) we would get a term with $Z(s_1,s_2)$, a term
with $Z(s_3,s_2)$ and a term with $Z(s'_3,s_2)$. This looks suspiciously
like a large-scale Bellman from $s_1$... Variance could be higher since
we don't use the model $Z$ and use sampling instead. To be tested.
On-policy, another
option is to sample triplets $(s_1,s_3,s_2)$ that are some time apart on
the trajectory, e.g., starting with $s_3$ uniform along the past
trajectory, and moving back for $s_1$ and forward for $s_2$ (with factors
$\gamma$). Looks a lot like eligibility traces?

>>>>

\paragraph{Better variant?}
I think
everything works equally well if we just learn a kernel $Z(s_1,s_2)$ such
that
\begin{equation}
\transp{\1_{s_1}}\Delta^{-1}\psi\approx \E_{s_2\sim \rho} \,Z(s_1,s_2)
\psi(s_2)
\end{equation}
and this is still linear in $\psi$. This represents the inverse Laplace
operator by
\begin{equation}
Z=\sum_{s_1,s_2}\rho(s_2)  Z(s_1,s_2)\1_{s_1}\transp{\1_{s_2}}
\end{equation}
In particular the value function is
\begin{equation}
V(s_1)=\E_{s_2\sim \rho} \, Z(s_1,s_2) R(s_2)
\end{equation}
Downside: this representation of the value function does not allow for first
computing a representation of $R$ as $\E_{s_2} R(s_2)f(s_2)$ online, then taking a dot product with
$\green$. Here we would have to learn an auxiliary network to imitate
$\E_{s_2\sim \rho} Z(s_1,s_2) R(s_2)$ (but this is a supervised learning
problem; still, it looks suspiciously like $TD(1)$ except maybe for the
$O(1/(1-\gamma))$ variance). Alternatively, policy gradient can just directly be obtained
via
\begin{equation}
\partial \pi(a|s)\E_{s'|s,a} \E_{s_2\sim \rho} \, Z(s',s_2)R(s_2)
\end{equation}
(this works even if $\rho$ is \emph{not} the invariant distribution). We
can apply this each time a new reward is observed at some point $s_2$. We
can even set another arbitrary distribution on $s_1$: this policy
gradient will maximize
the reward obtained from a starting point from that distribution.

@@

On-policy, sampling from $Z(\cdot,s_3)$ is easy: just sample one of the
recent states with geometric rates $\gamma$. This should divide the
right part of the equation by $Z(s_1,s_3)$ (up to some $\gamma$-dependent
factor). Thus we would sample two states $s_1,s_2$ along the trajectory,
sample a state $s_3$ in between on the trajectory from $s_1$, and (with
the right $\gamma$ factors) we would get a term with $Z(s_1,s_2)$, a term
with $Z(s_3,s_2)$ and a term with $Z(s'_3,s_2)$. This looks suspiciously
like a large-scale Bellman from $s_1$... Variance could be higher since
we don't use the model $Z$ and use sampling instead. To be tested. Write
carefully because $s'_3$ should be uncorrelated from $s_2$!! So not
possible with a single
trajectory?
On-policy, another
option is to sample triplets $(s_1,s_3,s_2)$ that are some time apart on
the trajectory, e.g., starting with $s_3$ uniform along the past
trajectory, and moving back for $s_1$ and forward for $s_2$ (with factors
$\gamma$). Looks a lot like eligibility traces?
